---
title: "Variola louti"
author: "Eva Schemmel"
date: "`r Sys.Date()`"
output: pdf_document

---

## Life history analysis for *Variola louti* from Guam

Code provided for the analyis of size/age at maturity, size/age at sex change, spawning season and lunar spawning frequency. Bayesian growth model and extracting SST from ERDDAP for comparision in edge analysis are in separate R files.

1.  Load packages, and load and clean data.

```{r setup, message=FALSE, warning=FALSE, include=TRUE}
library(dplyr)
library(reshape2)
library(lme4)
library(boot)
library(magrittr)
library(plotrix)
library(lubridate)
library(plyr)
library(car)
library(lunar)
library(mgcv)
library(itsadug)
library(ggplot2)
library(ggpubr)

#VALO Reproduction Data & Cleaning
setwd("~/Documents/research/VALO/Vlouti")
valo_r<-read.csv("Reproduction_4_6_2021.csv")

valo_r<-valo_r %>% subset(Species=="Variola louti") #N=273 
valo_r$ScientificName<-valo_r$Species

#load metadata and sampling data to link with reproduction information
setwd("~/Documents/github/TBP")
lhd<-read.csv("guam_7_22_2019.csv")

lhd$Date<-as.Date(lhd$FishedDate, format = "%m/%d/%Y")
lhd$Sex<-revalue(lhd$Sex, c("1 - Female"="F", "2 - Male"="M","3 - Unknown"="M"))


setwd("~/Documents/research/VALO/Vlouti")

lhd<-lhd%>% subset(ScientificName=="Variola louti")
lhd$SampleID=lhd$SampleTag
lhd<-lhd %>% select(SampleID,Length.cm., Weight.g., Sex, GonWeight, Date, Year, Month)

v_r<-left_join(valo_r, lhd)

#age data
age<-read.csv("Age_VALO_March_2023.csv")
valo_ar<-left_join(age, v_r, by="SampleID")

#sub data sets for plotting use later
valo_ar$H_Sex<-as.factor(valo_ar$H_Sex)
valo_ar_f<-valo_ar %>%
  subset(H_Sex=="F") #female
valo_ar_m<-valo_ar %>%
  subset(H_Sex=="M" ) #male
valo_ar_po<-valo_ar %>%
  subset(H_Sex=="M_PO" ) #male with primary oocytes
valo_ar_t<-valo_ar %>%
  subset(H_Sex=="T" ) #transitional

```

##Spawning Season

First assess spawning season, then can assess L50. Need to constrain data for L50 to within the spawning season. Since some spawning occurs all year for this species (spawning capable individuals found all year), all the data is used for determining L50. Also look at relationships with gonadosomatic index (GSI) by fish length and reproductive phase.

```{r  spawningseason,  message=FALSE, warning=FALSE, include=TRUE, fig.cap="Female Guam Variola louti gonadosomatic index (GSI) versus fork length, and spawning capable and actively spawning female log gonad weight (GW) versus gonad free body weight (GFBW) (log GW = 1.3860 + GFBW * 0.0016, df = 40, R2=0.23) for fish sampled from Guam between 2010 and 2017. Spawning capable and actively spawning females are black, immature females are white, and all other female reproductive phases are grey."}

gsi<- v_r%>% mutate(GSI=GonWeight/Weight.g.*100)
gsi$Month<-month(gsi$Date, label=TRUE) #get monthly labels for summarizing and plotting

#separate sexes for analysis
gsi_m<-subset(gsi, H_Sex=="M") #males
gsi<-subset(gsi, H_Sex=="F") #females
gsi<- subset(gsi, GSI>=0.00001) #clean and remove NA and zeros

###GSI means and sd by reproductive phase###
gsi<- gsi[complete.cases(gsi$Length.cm.),]  #subset data to get records that have Fork Lengths
gsi_phase<-gsi %>%
  dplyr::group_by(Phase) %>%
  dplyr::summarise(mean=mean(GSI), n=length(GSI), Sd=sd(GSI))
gsi_phase$SE <-gsi_phase$Sd/sqrt(gsi_phase$n) #adding standard error to the dataframe

gsi_phase_oocyte<-gsi %>%
  dplyr::group_by(Oocyte_Stage) %>%
  dplyr::summarise(mean=mean(GSI), n=length(GSI), Sd=sd(GSI))

#developing phase gsi and sd (VTI and VTII)
developing_gsi<-gsi %>% subset(Oocyte_Stage=="VTI" | Oocyte_Stage=="VTII")
developing_gsi<-developing_gsi %>%
  dplyr::summarise(mean=mean(GSI), n=length(GSI), Sd=sd(GSI))
developing_gsi$SE <-developing_gsi$Sd/sqrt(developing_gsi$n)

####GSI vs FL relationship#########
gsi<- gsi[complete.cases(gsi$GSI),] 
gsi <- gsi[order(gsi$Length.cm.),]
gsi_sc<-gsi %>% subset(Phase=="SC" |Phase=="AS")
gsi_sc$GFBW=gsi_sc$Weight.g.-gsi_sc$GonWeight

######FL and GSI plots####
{par(mfrow=c(1,2),mar=c(4,4,3,3))
plot(gsi$GSI~gsi$Length.cm., bty="l", ylab="GSI", xlab="Fork Length (cm)", pch=19, las=1,xaxs="i",yaxs="i", xlim=c(20,45), ylim=c(0,10))
gsiU<-gsi %>% subset(Phase=="U")
gsiD<-gsi %>% subset(Phase=="D"|Phase=="RG"|Phase=="RE")
gsiSC<-gsi %>% subset(Phase=="SC"|Phase=="U")
points(gsiU$GSI~gsiU$Length.cm., pch=21,col="black",bg="white", add=TRUE)
points(gsiD$GSI~gsiD$Length.cm., pch=21,col="black",bg="grey", add=TRUE)
text(21,9.5, "a")
plot(log(gsi_sc$GonWeight)~gsi_sc$GFBW, bty="l", ylab="log GW", xlab="GFBW", pch=19,xaxs="i",yaxs="i",xlim=c(100,1400), ylim=c(0,5))
axis(1, seq(200,1400,by=200), labels =seq(200,1400,by=200) )
expr<-lm(log(gsi_sc$GonWeight)~gsi_sc$GFBW)
leng<-seq(100,1200,by=10)
lines(gsi_sc$GFBW,(predict(expr, list(x=leng))), lwd=2)
text(260,4.8, "b")}

summary(expr)

```

```{r gsi,message=FALSE, warning=FALSE, include=TRUE, fig.cap="Median and interquartile range of gonadosomatic index (GSI) for mature female (a) and male (b) Variola louti by month" }

#####box plot spawning season#####
#female and sample sizes per month
guam_total<-as.data.frame(gsi %>%
                            dplyr::group_by(Month)%>%
                            dplyr::summarise(total=length(Length.cm.))) #female
male_total<-as.data.frame(gsi_m %>%
                            dplyr::group_by(Month)%>%
                            dplyr::summarise(total=length(Length.cm.))) #male


{par(mfrow=c(2,1),mar=c(3.5,3.5,1.5
                       ,1),mgp=c(2,.7,0))
boxplot(gsi$GSI~gsi$Month,bty="l",ylab="GSI",xlab="", ylim=c(0.0,10),frame=T,cex.axis=1.0, cex.lab=1.0,cex=1.0, cex.main=1.0, las=1)
par(xpd=TRUE)
text(x =guam_total$Month, y=11.2, label = guam_total$total , offset=0, cex = 1.0)#too add sample size above plot
text(0.3,9, "a", cex=1.0)
par(xpd=TRUE)
boxplot(gsi_m$GSI~gsi_m$Month,bty="l",ylab="GSI",xlab="", ylim=c(0,.5), frame=T,cex.axis=1.0, cex.lab=1.0,cex=1.0, cex.main=1.0, las=1)
text(x =male_total$Month, y=0.56, label = male_total$total , offset=0, cex = 1.0)#too add sample size above plot
text(0.3,.45, "b", cex=1.0)}

```

## Proportion of Female Reproductive Phases per Month

```{r reproductivephases, message=FALSE, warning=FALSE, include=TRUE, fig.cap = "Trend in reproductive stages of female Variola louti by month for fish sampled from Guam between 2010 and 2017. Developing stages (Developing I & Developing II) were grouped together into “Developing.” Numbers above each bar indicate aggregate monthly sample sizes."}

###female histo reproductive phase per month frequency graph###
#number of females in each reproductive phase
temp<-gsi
temp <- temp[complete.cases(temp$Length.cm.),] 
temp$PHASE<-factor(temp$Phase, levels=c( "AS","SC","D","RG","RE", "U"), ordered=TRUE)
tempAS<-subset(temp, PHASE=="AS")
tempSC<-subset(temp, PHASE=="SC")
tempD<-subset(temp, PHASE=="D")
tempRG<-subset(temp, PHASE=="RG")
tempRE<-subset(temp, PHASE=="RE")
tempU<-subset(temp, PHASE=="U")

#to compare the lengths of immature and regenerating females - this can be useful to look for outliers
#hist(tempRE$Length.cm.)
#hist(tempU$Length.cm.)

##percent of females that are immature 
fem_total<-length(temp$PHASE)
u=length(tempU$Phase)
immaturef<-u/fem_total 

prop_month_cast <- dcast(temp, Month~PHASE, value.var="PHASE")

prop_month_cast$sum <- rowSums(prop_month_cast[2:7])# make sure it adds to 100%
prop_month_cast$AS<-(prop_month_cast$AS/prop_month_cast$sum)*100
prop_month_cast$D<-(prop_month_cast$D/prop_month_cast$sum)*100
prop_month_cast$SC<-(prop_month_cast$SC/prop_month_cast$sum)*100
prop_month_cast$RE<-(prop_month_cast$RE/prop_month_cast$sum)*100
prop_month_cast$U<-(prop_month_cast$U/prop_month_cast$sum)*100
prop_month_cast$RG<-(prop_month_cast$RG/prop_month_cast$sum)*100

# ###add in missing months of sampling if there are any, example below
# #install.packages("tibble")
# prop_month_cast<-add_row(prop_month_cast, Month=6, SC=0, D=0, RE=0,RG=0, U=0, sum=0)

prop_month_cast<-prop_month_cast %>% arrange(Month)
prop_month_cast <- prop_month_cast[complete.cases(prop_month_cast$Month),] 

counts.L <- t(prop_month_cast[2:7])
count.labL <- c("Jan", "Feb", "Mar", "Apr", "May","Jun", "Jul","Aug","Sep","Oct","Nov","Dec")
sample_n=prop_month_cast$sum

{par(mfrow=c(1,1),mar=c(4,5,2,11))
sz<-barplot(counts.L,
            names.arg=count.labL, cex.lab=1.0, cex.axis=1.0,xaxs="i",yaxs="i",cex=1.0,ylab="Percent Frequency", main="", col=c( "firebrick", "tan3", "palegoldenrod","mediumpurple2", "lightcyan2", "steelblue"), las=1)
par(xpd=TRUE)
legend(15,80,bty="n",legend=c("Actively Spawning", "Spawning Capable","Developing","Regenerating","Regressing" ,"Undeveloped"),cex=1.1
       ,col=c( "firebrick", "tan3","palegoldenrod", "mediumpurple2","lightcyan2", "steelblue"),pch=15, border=TRUE) #make legend to plot outside the plotting space need to do new par space
text(x = sz, y=104, label = sample_n , offset=10, cex = 1.0)}#too add sample size above plot

```


## Lunar Spawning Assessment

```{r lunar, message=FALSE, warning=FALSE, include=TRUE, fig.cap="Generalized additive model GSI ~ Fork Length+s(Lunar Day) fit to raw data points (black line) by lunar day (3 day mean)."}
#####Lunar Spawning#####
#uses libraries lunar, mgcv, itsadug

#subsetting data for functionally mature females and males 
gsi<-gsi %>% subset(M_VT==1) #functionally mature females
gsi_m<-gsi_m %>% subset(Mature==1) #mature males

gsi$lunar.phase<-lunar::lunar.phase(gsi$Date,shift=-10, name=8)
lunar_total<-as.data.frame(gsi %>%
                            dplyr::group_by(lunar.phase)%>%
                            dplyr::summarise(total=length(Length.cm.)))

gsi <-
  gsi  %>%
  mutate(lunar_day = lunar.phase(as.Date(Date), shift=-10) / 0.212769) 
gsi$lunar_day<-round(gsi$lunar_day,0)

# GAM uses library mgcv
#testing two models and which fits the data better
mod1<-gam(log(GSI)~Length.cm.+s(lunar_day), data=gsi)
summary(mod1)
mod2<-gam(log(GSI)~s(lunar_day), data=gsi)
summary(mod2)
#To look at models graphically 
#plot.gam(mod1, shade=TRUE)
#plot.gam(mod2, shade=TRUE)

#use AIC to select best model
AIC(mod1, mod2)

#to create a box plot of GSI by lunar day need to bin GIS by lunar day - used 3 day bins
gsitemp=gsi
brks <- seq(from=0,to=30,by=3)
gsitemp$binl <- cut(gsitemp[,"lunar_day"], breaks=brks, labels=brks[1:(length(brks)-1)],right=F)
gsitemp$binl<-as.factor((gsitemp$binl))
gsitemp<-gsitemp %>%
  dplyr::arrange(desc(lunar_day))%>%
  dplyr::group_by(binl)%>%
  mutate(Mean=mean(GSI, na.rm=T),lower=quantile(GSI,probs = 0.025, na.rm=TRUE), upper=quantile(GSI,probs = 0.0975, na.rm=TRUE), upperbox=quantile(GSI,probs = 0.75,lowerbox=quantile(GSI,probs = 0.25, na.rm=TRUE))) #summarize or mutate
gsitemp$binl<-as.numeric(as.character((gsitemp$binl)))

gsilunarsum<-gsitemp %>%
  dplyr::group_by(binl) %>%
  dplyr::summarise(total=n())
gsilunarsum<-gsilunarsum[1:10,]
gsilunarsum$lunar_day<-gsilunarsum$binl

#make shift code to line up box plots on the x axis over the median lunar day for the bin
c(0,3,6,9,12,15,18,21,24,27)
gsitemp1<-gsitemp%>%
 ungroup()%>%
  select(GSI, binl)
l0<-gsitemp1%>%
  subset(binl==0)
l0<-l0 %>%
  select(GSI)
l3<-gsitemp1%>%
  subset(binl==3)
l6<-gsitemp1%>%
  subset(binl==6)
l9<-gsitemp1%>%
  subset(binl==9)
l12<-gsitemp1%>%
  subset(binl==12)
l15<-gsitemp1%>%
  subset(binl==15)
l18<-gsitemp1%>%
  subset(binl==18)
l21<-gsitemp1%>%
  subset(binl==21)
l24<-gsitemp1%>%
  subset(binl==24)
l27<-gsitemp1%>%
  subset(binl==27)
xplace=as.numeric(c(0,3,6,9,12,15,18,21,24,27))

#lunar spawning figure#
{plot(gsitemp$Mean~gsitemp$lunar_day, ylab="GSI", xlab="Lunar Day", bty="n",frame=T,cex.axis=1.0,ylim=c(0,6),xaxs="i", cex.lab=1.0,cex=1.0, cex.main=1.0, las=1, pch=19, col="white")
#axis(1,at=c(seq(0,30,by=3)), cex.lab=1.0,cex.axis=1.0, las=1)
boxplot(l0$GSI, l3$GSI, l6$GSI, l9$GSI, l12$GSI,l15$GSI,l18$GSI,l21$GSI,l24$GSI,l27$GSI,
        at = c(1.5,4.5,7.5,10.5,13.5,16.5,19.5,22.5,25.5,28.5),
        names = c("0", "3", "6", "9","12","15","18","21","24","27"),
        las = 2,
        boxwex=3, add=TRUE, bty='n', xaxt="n")
plot_smooth(mod1, view="lunar_day", lwd=2,
            transform = exp, se=FALSE,rug=FALSE, plot_all=NULL,bty='n',add=TRUE) }#plot best fit gam model fit to all the data

```


## Size/age at Maturity 


```{r L50/A50, message=FALSE, warning=FALSE, include=TRUE}

#L50 functions
# setup functions ---------------------------------------------------------
data.setup <- function(dataframe,xname,yname){
  templ <- dataframe
  brks <- seq(from=min(templ[,xname]),to=max(templ[,xname]),by=2)
  templ$binl <- cut(templ[,xname], breaks=brks, labels=brks[1:(length(brks)-1)],right=F)
  templ <- dcast(templ, binl~templ[,yname])
  colnames(templ) <- c("binl","a","b")
  templ$binl <- as.character(templ$binl)
  templ$binl <- as.numeric(templ$binl)
  templ$sum <- rowSums(templ[2:3])
  templ$a.p <- templ$a/templ$sum
  templ$b.p <- templ$b/templ$sum
  return(templ)
}

data.setup <- function(dataframe,xname,yname){
  templ <- dataframe
  brks <- seq(from=20,to=50,by=2)
  templ$binl <- cut(templ[,xname], breaks=brks, labels=brks[1:(length(brks)-1)],right=F)
  templ <- dcast(templ, binl~templ[,yname])
  colnames(templ) <- c("binl","a","b")
  templ$binl <- as.character(templ$binl)
  templ$binl <- as.numeric(templ$binl)
  templ$sum <- rowSums(templ[2:3])
  templ$a.p <- templ$a/templ$sum
  templ$b.p <- templ$b/templ$sum
  return(templ)
}

data.setup_age <- function(dataframe,xname,yname){
  templ <- dataframe
  brks <- seq(from=min(templ[,xname]),to=max(templ[,xname]),by=1)
  templ$binl <- cut(templ[,xname], breaks=brks, labels=brks[1:(length(brks)-1)],right=F)
  templ <- dcast(templ, binl~templ[,yname])
  colnames(templ) <- c("binl","a","b")
  templ$binl <- as.character(templ$binl)
  templ$binl <- as.numeric(templ$binl)
  templ$sum <- rowSums(templ[2:3])
  templ$a.p <- templ$a/templ$sum
  templ$b.p <- templ$b/templ$sum
  return(templ)
}


model.ci <- function(dataframe,xname,yname){
  templ <- dataframe
  glm.out <- glm(templ[,yname] ~ templ[,xname], family=binomial(link=logit), data=templ)
  newdata.l=data.frame(templ[,xname])
  colnames(newdata.l) <- xname
  templ <- predict(glm.out,newdata=newdata.l, type="response", se.fit=T)
  ci <- matrix(c(templ$fit+1.96*templ$se.fit,templ$fit-1.96*templ$se.fit),ncol=2)
  ci <- data.frame(ci)
  ci <- cbind(ci,newdata.l)
  ci <- ci[with(ci,order(ci[,3])),]
  return(ci)
}

est <- function(coef,p) (log(p/(1-p))-coef[1])/coef[2] #p is the proportion

boot.est <- function(dataframe,xname,yname){
  myfun <- function(dataframe,xname,yname){
    templ <- dataframe
    srows <- sample(1:nrow(templ),nrow(templ),TRUE)
    glm.out <- glm(templ[,yname] ~ templ[,xname], family=binomial(link=logit), data=templ[srows,])
    return(est(coef(glm.out),0.5))
  }
  bootdist <- replicate(1000, myfun(dataframe,xname,yname)) 
  boot.ci <- quantile(unlist(bootdist),c(.025,.975)) # 95% CI
  return(boot.ci)
}


repro_f<-subset(v_r, H_Sex=="F")
#repro_f<-subset(repro_f, Phase!="RE")
repro_f <- repro_f[complete.cases(repro_f$Length.cm.),] 
#repro_f<-repro_f %>% subset(Month=="Apr"|Month=="May"| Month=="Jun" |Month=="Jul" | Month=="Aug"|Month=="Sep")
lbins.l50vt <- data.setup(repro_f,"Length.cm.","M_VT") 
glm.l50vt <- glm(M_VT ~ Length.cm., family=binomial(link=logit), data=repro_f)
l50.civt <- model.ci(repro_f,"Length.cm.","M_VT")

length(repro_f$Length.cm.)
bcL_tlvt<-bootCase(glm.l50vt, R=1000)
lrPerc <- function(cf,p) (log(p/(1-p))-cf[1])/cf[2]#p is the proportion
l50vt<-lrPerc(coef(glm.l50vt), 0.5)
l50vt<-apply(bcL_tlvt, 1, lrPerc, p=0.5)
l50civt<-quantile(l50vt, c(0.25,0.975))
l50vt <- est(coef(glm.l50vt),.5)

#physiological maturity
lbins.l50ca <- data.setup(repro_f,"Length.cm.","M_CA") 
glm.l50ca <- glm(M_CA ~ Length.cm., family=binomial(link=logit), data=repro_f)
l50.cica <- model.ci(repro_f,"Length.cm.","M_CA")

length(repro_f$Length.cm.)
bcL_tlca<-bootCase(glm.l50ca, R=1000)
lrPerc <- function(cf,p) (log(p/(1-p))-cf[1])/cf[2]#p is the proportion
l50ca<-lrPerc(coef(glm.l50ca), 0.5)
l50ca<-apply(bcL_tlca, 1, lrPerc, p=0.5)
l50cica<-quantile(l50ca, c(0.25,0.975))
l50ca <- est(coef(glm.l50ca),.5)

######Size at functional maturity####
plot(lbins.l50vt$bin,lbins.l50vt$b.p,pch=19,cex.axis=1.0, cex=1.0,cex.lab=1.0, bty="l",xlab="Fork Length (cm)",ylab="Percent Mature"
     ,ylim=c(-0.1,1.1),yaxt="n",yaxs="i",xaxs="i", xlim=c(16,40))
axis(2,at=c(seq(0,1.1,by=.1)),labels=c(seq(0,110,by=10)), cex.lab=1.0,cex.axis=1.0, las=1)
curve((exp(coef(glm.l50vt)[1]+coef(glm.l50vt)[2]*x))/(1+exp(coef(glm.l50vt)[1]+coef(glm.l50vt)[2]*x)),add=T)
lines(l50.civt[,3],l50.civt[,1],lty=2)
lines(l50.civt[,3],l50.civt[,2],lty=2)
lines(c(l50vt,l50vt),c(-0.1,.5), lwd=2,col="grey",lty=1, font=2)
lines(c(16,l50vt),c(.5,.5), lwd=2,col="grey",lty=1,font=2)

#age at maturity
valo_ar_f<-valo_ar %>%
  subset(H_Sex=="F")

valo_ar_f<-valo_ar_f[complete.cases(valo_ar_f$reage),] #make sure there are no blanks in ages 
abins.a50vt <- data.setup_age(valo_ar_f,"reage","M_VT") 
glm.a50vt <- glm(M_VT ~ reage, family=binomial(link=logit), data=valo_ar_f)
a50.civt <- model.ci(valo_ar_f,"reage","M_VT")
coef(glm.a50vt)
length(valo_ar_f$reage)
bcL_tlvt<-bootCase(glm.a50vt, R=1000)
lrPerc <- function(cf,p) (log(p/(1-p))-cf[1])/cf[2]#p is the proportion
a50vt<-lrPerc(coef(glm.a50vt), 0.5)
a50vt<-apply(bcL_tlvt, 1, lrPerc, p=0.5)
a50civt<-quantile(a50vt, c(0.25,0.975))
a50vt <- est(coef(glm.a50vt),.5)

#physiological
abins.a50ca <- data.setup_age(valo_ar_f,"reage","M_CA") 
glm.a50ca <- glm(M_CA ~ reage, family=binomial(link=logit), data=valo_ar_f)
a50.cica <- model.ci(valo_ar_f,"reage","M_CA")

length(valo_ar_f$reage)
bcL_tlca<-bootCase(glm.a50ca, R=1000)
lrPerc <- function(cf,p) (log(p/(1-p))-cf[1])/cf[2]#p is the proportion
a50ca<-lrPerc(coef(glm.a50ca), 0.5)
a50ca<-apply(bcL_tlca, 1, lrPerc, p=0.5)
a50cica<-quantile(a50ca, c(0.25,0.975))
a50ca <- est(coef(glm.a50ca),.5)


#####Size at sex change######
v_r$SexID<-v_r$H_Sex
v_r$SexID<-revalue(v_r$SexID, c("F"=0, "T"=1, "M"=1,"M_PO"=1))
v_r$SexID<-as.factor(v_r$SexID)
v_r<-v_r[complete.cases(v_r$Length.cm.),]

#remove rows with blanks
v_r<-v_r[!(is.na(v_r$SexID) | v_r$SexID==""), ] #removed two rows with blank H_Sex
lbins.l50sex <- data.setup(v_r,"Length.cm.","SexID") 
glm.l50sex <- glm(SexID ~ Length.cm., family=binomial(link=logit), data=v_r)
l50.cisex <- model.ci(v_r,"Length.cm.","SexID")

bcL_tlsex<-bootCase(glm.l50sex, R=1000)
lrPerc <- function(cf,p) (log(p/(1-p))-cf[1])/cf[2]#p is the proportion
l50sex<-lrPerc(coef(glm.l50sex), 0.5)
l50sex<-apply(bcL_tlsex, 1, lrPerc, p=0.5)
l50cisex<-quantile(l50sex, c(0.25,0.975))
l50sex <- est(coef(glm.l50sex),.5)

{plot(lbins.l50sex$bin,lbins.l50sex$b.p,pch=19,cex.axis=1.0, cex=1.0,cex.lab=1.0, bty="l",xlab="Fork Length (cm)",ylab="Percent Mature"
     ,ylim=c(-0.1,1.1),yaxt="n",yaxs="i",xaxs="i", xlim=c(16,50))
axis(2,at=c(seq(0,1.1,by=.1)),labels=c(seq(0,110,by=10)), cex.lab=1.0,cex.axis=1.0, las=1)
curve((exp(coef(glm.l50sex)[1]+coef(glm.l50sex)[2]*x))/(1+exp(coef(glm.l50sex)[1]+coef(glm.l50sex)[2]*x)),add=T)
lines(l50.cisex[,3],l50.cisex[,1],lty=2)
lines(l50.cisex[,3],l50.cisex[,2],lty=2)
lines(c(l50sex,l50sex),c(-0.1,.5), lwd=2,col="grey",lty=1, font=2)
lines(c(16,l50sex),c(.5,.5), lwd=2,col="grey",lty=1,font=2)}

#Age at sex change
valo_ar$SexID<-valo_ar$H_Sex
valo_ar$SexID<-revalue(valo_ar$SexID, c("F"=0, "T"=1, "M"=1,"M_PO"=1))
valo_ar$SexID<-as.factor(valo_ar$SexID)
valo_ar<-valo_ar[complete.cases(valo_ar$SexID),] 
valo_ar<-valo_ar[complete.cases(valo_ar$reage),] 
valo_ar<-valo_ar[complete.cases(valo_ar$Length.cm.),] ###went from 273 to 257
valo_ar<-valo_ar[!(is.na(valo_ar$SexID) | valo_ar$SexID==""), ] #removed two rows with blank H_Sex

abins.a50sex <- data.setup_age(valo_ar,"reage","SexID") 
abins.a50sex[12,1]<-14 #not sure why it dropped the last age
glm.a50sex <- glm(SexID ~ reage, family=binomial(link=logit), data=valo_ar)
a50.cisex <- model.ci(valo_ar,"reage","SexID")

length(valo_ar$Length.cm.)
bcL_tlsex<-bootCase(glm.a50sex, R=1000)
lrPerc <- function(cf,p) (log(p/(1-p))-cf[1])/cf[2]#p is the proportion
a50sex<-lrPerc(coef(glm.a50sex), 0.5)
a50sex<-apply(bcL_tlsex, 1, lrPerc, p=0.5)
a50cisex<-quantile(a50sex, c(0.25,0.975))
a50sex <- est(coef(glm.a50sex),.5)

#Age sex change
plot(abins.a50sex$bin,abins.a50sex$b.p,pch=19,cex.axis=1.0, cex=1.0,cex.lab=1.0, bty="l",xlab="Age (years)",ylab="Percent Mature"
     ,ylim=c(-0.1,1.1),yaxt="n",yaxs="i",xaxs="i", xlim=c(0,14))
axis(2,at=c(seq(0,1.1,by=.1)),labels=c(seq(0,110,by=10)), cex.lab=1.0,cex.axis=1.0, las=1)
curve((exp(coef(glm.a50sex)[1]+coef(glm.a50sex)[2]*x))/(1+exp(coef(glm.a50sex)[1]+coef(glm.a50sex)[2]*x)),add=T)
lines(a50.cisex[,3],a50.cisex[,1],lty=2)
lines(a50.cisex[,3],a50.cisex[,2],lty=2)
lines(c(a50sex,a50sex),c(-0.1,.5), lwd=2,col="grey",lty=1, font=2)
lines(c(0,a50sex),c(.5,.5), lwd=2,col="grey",lty=1,font=2)


```

## Bubble Plot for Size/Age at Maturity and Size/Age at Sex Chnage

```{r bubble,  message=FALSE, warning=FALSE, include=TRUE}
#length at maturity
valo_f<-valo_ar %>%
  subset(H_Sex=="F")
l50bubble<- ggplot()+
  geom_smooth(data = valo_f, aes(x = Length.cm., y = M_VT),
              method = "glm", method.args = list(family = "binomial"), 
              se = FALSE, color="black") +
  geom_point(data=lbins.l50vt, aes(x = binl, y =b.p, size=sum), alpha=0.5) +
  scale_size(range = c(.5, 5), name="Sample Size")+
  geom_line(aes(x=l50.civt[,3],y=l50.civt[,1], group=1), linetype="dashed", size= 0.75)+ 
  geom_line(aes(x=l50.civt[,3],y=l50.civt[,2], group=1), linetype="dashed", size= 0.75)+ 
  geom_segment(aes(x = 19.9, y = 0.5, xend =l50vt , yend = 0.5))+ #to add L50 lines if wanted
  geom_segment(aes(x = l50vt, y = -0.01, xend =l50vt , yend = 0.5))+ #to add L50 lines if wanted
  theme_bw() +
  theme(
    panel.border = element_blank(),
    panel.grid = element_blank(),
    axis.title = element_text(size = 12),
    axis.title.y = element_text(margin= margin(r= 15)), # move title away from axis,
    axis.title.x = element_text(margin= margin(t= 15)),
    legend.position = "none",
   # legend.text = element_text(colour="black", size=10, face="bold"),
    #legend.title = element_text(colour="black", size=10, face="bold"),
    axis.text = element_text(colour = "black", size = 12),
    axis.line = element_line(colour = "black")) +
  xlab('Fork Length (cm)') +
  ylab("Proportion Mature")+
  scale_x_continuous(name = "Fork Length (cm)",breaks=seq(20,45,5)) 

#l50bubble<-l50bubble+theme(legend.position = c(.87,.2),
#legend.text = element_text(colour="black", size=12),
#legend.title = element_text(colour="black", size=12))

l50bubble<- l50bubble+ scale_y_continuous(expand = c(0, 0), limits = c(-0.01, 1.02))+
  scale_x_continuous(expand = c(0, 0), limits = c(19.9,45), breaks=seq(20,45,5))


#Age at maturity
 a50bubble<- ggplot()+
  geom_smooth(data = valo_f, aes(x = reage, y = M_VT),
              method = "glm", method.args = list(family = "binomial"), 
              se = FALSE, color="black") +
  geom_point(data=abins.a50vt, aes(x = binl, y =b.p, size=sum), alpha=0.5) +
  scale_size(range = c(.5, 5), name="Sample Size")+
  geom_line(aes(x=a50.civt[,3],y=a50.civt[,1], group=1), linetype="dashed", size= 0.75)+ 
  geom_line(aes(x=a50.civt[,3],y=a50.civt[,2], group=1), linetype="dashed", size= 0.75)+ 
  geom_segment(aes(x = 0, y = 0.5, xend =a50vt , yend = 0.5))+ #to add L50 lines if wanted
  geom_segment(aes(x = a50vt, y = -0.01, xend =a50vt , yend = 0.5))+ #to add L50 lines if wanted
  theme_bw() +
  theme(
    panel.border = element_blank(),
    panel.grid = element_blank(),
    axis.title = element_text(size = 12),
    axis.title.y = element_text(margin= margin(r= 15)), # move title away from axis,
    axis.title.x = element_text(margin= margin(t= 15)),
    legend.position = "none",
    legend.text = element_blank(),
    legend.title = element_blank(),
    axis.text = element_text(colour = "black", size = 12),
    axis.line = element_line(colour = "black")) +
  xlab('Age (Years)') +
  ylab("Proportion Mature")+
  scale_x_continuous(name = "Age (years)",breaks=seq(0,14,2)) 


  a50bubble<- a50bubble+ scale_y_continuous(expand = c(0, 0), limits = c(-0.01, 1.02))+
    scale_x_continuous(expand = c(0, 0), limits = c(0, 12), breaks=seq(0,12,2))
  
  
  
##length at sex change
  l50sexbubble<- ggplot()+
    geom_smooth(data = v_r, aes(x = Length.cm., y = as.numeric(as.character(SexID))),
                method = "glm", method.args = list(family = "binomial"), 
                se = FALSE, color="black") +
    geom_point(data=lbins.l50sex, aes(x = binl, y =b.p, size=sum), alpha=0.5) +
    scale_size(range = c(.5, 5), name="Sample Size")+
    geom_line(aes(x=l50.cisex[,3],y=l50.cisex[,1], group=1), linetype="dashed", size= 0.75)+ 
    geom_line(aes(x=l50.cisex[,3],y=l50.cisex[,2], group=1), linetype="dashed", size= 0.75)+ 
    geom_segment(aes(x = 19.9, y = 0.5, xend =l50sex , yend = 0.5))+ #to add L50 lines if wanted
    geom_segment(aes(x = l50sex, y = -0.01, xend =l50sex , yend = 0.5))+ #to add L50 lines if wanted
    theme_bw() +
    theme(
      panel.border = element_blank(),
      panel.grid = element_blank(),
      axis.title = element_text(size = 12),
      axis.title.y = element_text(margin= margin(r= 15)), # move title away from axis,
      axis.title.x = element_text(margin= margin(t= 15)),
      legend.position = "none",
      legend.text = element_blank(),
      legend.title = element_blank(),
      axis.text = element_text(colour = "black", size = 12),
      axis.line = element_line(colour = "black")) +
    xlab('Fork Length (cm)') +
    ylab("Proportion Male")+
    scale_x_continuous(name = "Fork Length (cm)",breaks=seq(20,50,5)) 
  
  l50sexbubble<- l50sexbubble+ scale_y_continuous(expand = c(0, 0), limits = c(-0.01, 1.02), breaks=seq(0,1,.25))+
    scale_x_continuous(expand = c(0, 0), limits = c(19.9,50), breaks=seq(20,50,5))
  
  
#age at sex chage
  valo_ar <- valo_ar[!(valo_ar$SexID == ""), ]
  valo_ar$SexID[valo_ar$SexID==' ']=NA
  valo_ar$SexID= droplevels(valo_ar$SexID)
  a50sexbubble<- ggplot()+
    geom_smooth(data = valo_ar, aes(x = reage, y = as.numeric(as.character(SexID))),
                method = "glm", method.args = list(family = "binomial"), 
                se = FALSE, color="black") +
       geom_point(data=abins.a50sex, aes(x = binl, y =b.p, size=sum), alpha=0.5) +
    scale_size(range = c(.5, 5), name=" ")+
    geom_line(aes(x=a50.cisex[,3],y=a50.cisex[,1], group=1), linetype="dashed", size= 0.75)+ 
    geom_line(aes(x=a50.cisex[,3],y=a50.cisex[,2], group=1), linetype="dashed", size= 0.75)+ 
    geom_segment(aes(x = 0, y = 0.5, xend =a50sex , yend = 0.5))+ #to add L50 lines if wanted
    geom_segment(aes(x = a50sex, y = -0.01, xend =a50sex , yend = 0.5))+ #to add L50 lines if wanted
    theme_bw() +
    theme(
      panel.border = element_blank(),
      panel.grid = element_blank(),
      axis.title = element_text(size = 12),
      axis.title.y = element_text(margin= margin(r= 15)), # move title away from axis,
      axis.title.x = element_text(margin= margin(t= 15)),
      legend.position = "none",
      legend.text = element_blank(),
      legend.title = element_blank(),
      axis.text = element_text(colour = "black", size = 12),
      axis.line = element_line(colour = "black")) +
    xlab('Age (Years)') +
    ylab("Proportion Male")+
    scale_x_continuous(name = "Age (years)",breaks=seq(0,18,2)) 
  
  
  a50sexbubble<-a50sexbubble + 
  scale_y_continuous(expand = c(0, 0), limits = c(-0.01, 1.02))+
    scale_x_continuous(expand = c(0, 0), limits = c(0, 18),breaks=seq(0,18,2))
  
  
  scale_y_continuous( limits = c(0.00, 1))

  #l50bubble<-l50bubble+theme(legend.position = c(.8,.25),
  #                           legend.text = element_text(colour="black", size=10),
  #                           legend.title = element_text(colour="black", size=10))
  
  a50sexbubble<- a50sexbubble+theme(legend.position = c(.87,.35),
                                legend.text = element_text(colour="black", size=12),
                                legend.title = element_text(colour="black", size=12))
  
  

  figure <- ggpubr::ggarrange(l50bubble, a50bubble,l50sexbubble,  a50sexbubble,
                      labels = c("", "", "",""),
                      ncol = 2, nrow = 2, 
                      hjust=-8.2,
                      vjust=0.6,
                      font.label = list(size = 12, color = "black", face = "plain", family = NULL))
  
  figure

```

## Length/Age Frequency Distribution

```{r distribution, message=FALSE, warning=FALSE, include=TRUE}
###age freq distribution#########
valo_f_u<-valo_ar %>%
  subset(H_Sex=="F" & Phase=="U")
valo_f_d<-valo_ar %>%
  subset(H_Sex=="F" & Phase=="D")
valo_f_sc<-valo_ar %>%
  subset(H_Sex=="F" & Phase=="SC")
valo_f_rg<-valo_ar %>%
  subset(H_Sex=="F" & Phase=="RG")
valo_f_re<-valo_ar %>%
  subset(H_Sex=="F" & Phase=="RE")
valo_t<-valo_ar %>%
  subset(H_Sex=="T")
valo_t_po<-valo_ar %>%
  subset(H_Sex=="M_PO")
valo_m<-valo_ar %>%
  subset(H_Sex=="M")

#Plot it. I save it width=2900, height=3800, res=250 t see all the details in the figure that you cant see here.
{par(mfcol=c(8,2), mar=c(1,5,2,1), oma=c(4,3,.4,.4))
hist(valo_f_u$Length.cm., right=FALSE,main="" ,xaxs="i",yaxs="i", breaks=seq(20,50,2.5), ylab="", xlab="" ,ylim=c(0,20), xaxt="n",las=1 , col="grey", cex=2, cex.lab=2, cex.axis=2)
axis(1,at= seq(20,50,by=5) , labels = seq(20,50,by=5) ,cex=2, cex.lab=2, cex.axis=2)
box(bty = "l")
text(20,18.5, "Undeveloped",cex=2, cex.lab=2, cex.axis=2, pos = 4)
hist(valo_f_d$Length.cm., right=FALSE,main="" ,xaxs="i",yaxs="i",breaks=seq(20,50,2.5), ylab="",xlab="" ,ylim=c(0,20),xaxt="n", las=1 , col="grey",cex=2, cex.lab=2, cex.axis=2)
axis(1,at= seq(20,50,by=5) , labels = seq(20,50,by=5),cex=2, cex.lab=2, cex.axis=2 )
box(bty = "l")
text(20,18.5, "Developing",cex=2, cex.lab=2, cex.axis=2, pos = 4)
hist(valo_f_sc$Length.cm., right=FALSE,main="" ,xaxs="i",yaxs="i",breaks=seq(20,50,2.5), ylab="",xlab="" ,ylim=c(0,20),xaxt="n", las=1 , col="grey", cex=2, cex.lab=2, cex.axis=2)
axis(1,at= seq(20,50,by=5) , labels = seq(20,50,by=5),cex=2, cex.lab=2, cex.axis=2  )
text(20,18.5, "Spawning Capable",cex=2, cex.lab=2, cex.axis=2, pos = 4)
box(bty = "l")
hist(valo_f_rg$Length.cm.,right=FALSE, main="" ,xaxs="i",yaxs="i",breaks=seq(20,50,2.5), ylab="", xlab="" ,ylim=c(0,20),xaxt="n",las=1 , col="grey", cex=2, cex.lab=2, cex.axis=2)
axis(1,at= seq(20,50,by=5) , labels = seq(20,50,by=5),cex=2, cex.lab=2, cex.axis=2  )
text(20,18.5, "Regressing",cex=2, cex.lab=2, cex.axis=2, pos = 4)
box(bty = "l")
hist(valo_f_re$Length.cm.,right=FALSE, main="" ,xaxs="i",yaxs="i",breaks=seq(20,50,2.5), ylab="", xlab="" ,ylim=c(0,20),xaxt="n",las=1 , col="grey", cex=2, cex.lab=2, cex.axis=2)
axis(1,at= seq(20,50,by=5) , labels = seq(20,50,by=5),cex=2, cex.lab=2, cex.axis=2  )
text(20,18.5, "Regenerating",cex=2, cex.lab=2, cex.axis=2,pos = 4)
box(bty = "l")
hist(valo_t$Length.cm.,right=FALSE, main="" ,xaxs="i",yaxs="i",breaks=seq(20,50,2.5), ylab="", xlab="" ,ylim=c(0,20),xaxt="n",las=1 , col="grey",cex=2, cex.lab=2, cex.axis=2)
axis(1,at= seq(20,50,by=5) , labels = seq(20,50,by=5) ,cex=2, cex.lab=2, cex.axis=2 )
text(20,18.5, "Transitional",cex=2, cex.lab=2, cex.axis=2, pos = 4)
box(bty = "l")
hist(valo_t_po$Length.cm.,right=FALSE, main="" ,xaxs="i",yaxs="i",breaks=seq(20,50,2.5), ylab="", xlab="" ,ylim=c(0,20),xaxt="n",las=1 , col="grey",cex=2, cex.lab=2, cex.axis=2)
axis(1,at= seq(20,50,by=5) , labels = seq(20,50,by=5),cex=2, cex.lab=2, cex.axis=2  )
text(20,18.5, "Male with PO",cex=2, cex.lab=2, cex.axis=2,pos = 4)
box(bty = "l")
hist(valo_m$Length.cm.,right=FALSE, main="" ,xaxs="i",yaxs="i",breaks=seq(20,50,2.5), ylab="", xlab="Fork Length (cm)" ,ylim=c(0,20),xaxt="n",las=1 , col="grey",cex=2, cex.lab=2, cex.axis=2)
axis(1,at= seq(20,50,by=5) , labels = seq(20,50,by=5), cex=2, cex.lab=2, cex.axis=2  )
text(20,18.5, "Male", cex=2, cex.lab=2, cex.axis=2,pos = 4)
box(bty = "l")
mtext(text="Fork Length (cm)",side=1,line=2,adj=0.22,outer=TRUE, cex=1.4)
mtext(text="Frequency",side=2,line=0,outer=TRUE, cex=1.4)

#age
hist(valo_f_u$reage, right=FALSE, main="" ,xaxs="i",yaxs="i", breaks=seq(0,18, 1), ylab="", xlab="" ,ylim=c(0,20), xaxt="n",las=1 , col="grey", cex=2, cex.lab=2, cex.axis=2)
axis(1,at= seq(0,18,by=2) , labels = seq(0,18,by=2) ,cex=2, cex.lab=2, cex.axis=2)
text(0.2,18.5, "Undeveloped",cex=2, cex.lab=2, cex.axis=2,pos = 4)
box(bty = "l")
hist(valo_f_d$reage,right=FALSE, main="" ,xaxs="i",yaxs="i",breaks=seq(0,18, 1), ylab="",xlab="" ,ylim=c(0,20),xaxt="n", las=1 , col="grey",cex=2, cex.lab=2, cex.axis=2)
axis(1,at= seq(0,18,by=2) , labels = seq(0,18,by=2),cex=2, cex.lab=2, cex.axis=2 )
text(0.2,18.5, "Developing",cex=2, cex.lab=2, cex.axis=2,pos = 4)
box(bty = "l")
hist(valo_f_sc$reage, right=FALSE,main="" ,xaxs="i",yaxs="i",breaks=seq(0,18, 1), ylab="",xlab="" ,ylim=c(0,20),xaxt="n", las=1 , col="grey", cex=2, cex.lab=2, cex.axis=2)
axis(1,at= seq(0,18,by=2) , labels = seq(0,18,by=2),cex=2, cex.lab=2, cex.axis=2  )
text(0.2,18.5, "Spawning Capable",cex=2, cex.lab=2, cex.axis=2,pos = 4)
box(bty = "l")
hist(valo_f_rg$reage,right=FALSE, main="" ,xaxs="i",yaxs="i",breaks=seq(0,18, 1), ylab="", xlab="" ,ylim=c(0,20),xaxt="n",las=1 , col="grey", cex=2, cex.lab=2, cex.axis=2)
axis(1,at= seq(0,18,by=2) , labels = seq(0,18,by=2),cex=2, cex.lab=2, cex.axis=2  )
text(0.2,18.5, "Regressing",cex=2, cex.lab=2, cex.axis=2,pos = 4)
box(bty = "l")
hist(valo_f_re$reage,right=FALSE, main="" ,xaxs="i",yaxs="i",breaks=seq(0,18, 1), ylab="", xlab="" ,ylim=c(0,20),xaxt="n",las=1 , col="grey", cex=2, cex.lab=2, cex.axis=2)
axis(1,at= seq(0,18,by=2) , labels = seq(0,18,by=2),cex=2, cex.lab=2, cex.axis=2  )
text(0.2,18.5, "Regenerating",cex=2, cex.lab=2, cex.axis=2,pos = 4)
box(bty = "l")
hist(valo_t$reage,right=FALSE, main="" ,xaxs="i",yaxs="i",breaks=seq(0,18, 1), ylab="", xlab="" ,ylim=c(0,20),xaxt="n",las=1 , col="grey",cex=2, cex.lab=2, cex.axis=2)
axis(1,at= seq(0,18,by=2) , labels = seq(0,18,by=2) ,cex=2, cex.lab=2, cex.axis=2 )
text(0.2,18.5, "Transitional",cex=2, cex.lab=2, cex.axis=2,pos = 4)
box(bty = "l")
hist(valo_t_po$reage,right=FALSE, main="" ,xaxs="i",yaxs="i",breaks=seq(0,18, 1), ylab="", xlab="" ,ylim=c(0,20),xaxt="n",las=1 , col="grey",cex=2, cex.lab=2, cex.axis=2)
axis(1,at= seq(0,18,by=2) , labels = seq(0,18,by=2),cex=2, cex.lab=2, cex.axis=2  )
text(0.2,18.5, "Male with PO",cex=2, cex.lab=2, cex.axis=2,pos = 4)
box(bty = "l")
hist(valo_m$reage,right=FALSE, main="" ,xaxs="i",yaxs="i",breaks=seq(0,18,by=1), ylab="", xlab="Age (yrs)" ,ylim=c(0,20),xaxt="n",las=1 , col="grey",cex=2, cex.lab=2, cex.axis=2)
axis(1,at= seq(0,18,by=2) , labels = seq(0,18,by=2), cex=2, cex.lab=2, cex.axis=2  )
text(0.2,18.5, "Male", cex=2, cex.lab=2, cex.axis=2,pos = 4)
box(bty = "l")
mtext(text="Age (yrs)",side=1,line=2, adj=.8,outer=TRUE, cex=1.4)}


```


## Male and female size and age differences

```{r sex statistics, message=FALSE, warning=FALSE, include=TRUE}
males<-rbind(valo_ar_m, valo_ar_po)
ks.test(valo_ar_f$Length.cm.,males$Length.cm.,   alternative = "greater")
ks.test(valo_ar_f$reage,males$reage,   alternative = "greater")

t.test(valo_ar_f$Length.cm., males$Length.cm.)
t.test(valo_ar_f$reage, males$reage)
max(males$reage, na.rm=TRUE)
mean(males$reage, na.rm=TRUE)
mean(valo_ar_f$reage, na.rm=TRUE)


```

